---
title: "SeaLevelRise_shiny"
author: "Alan Jackson"
format: dashboard
server: shiny
editor: source
---


```{r}
#| context: setup
### #| include: false

library(tidyverse)
library(stringr)
# library(tmap)
library(leaflet)

shiny::devmode(TRUE)

googlecrs <- "EPSG:4326"

# path <- "/home/ajackson/Dropbox/Rprojects/TexasDioceseCreationCare/Data/"
# Curated_path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/"
Curated_path <- "/home/ajackson/Dropbox/Rprojects/ERD/SeaLevel/"

Google_notes <- "https://docs.google.com/document/d/1ZYLhbcxoM2DShvo-eewVF4UHyNHtJVhjPB6KRr6JFI4/edit?usp=sharing"

Local_test <- TRUE

#   Directory where data is stored

DataLocation <- "https://www.ajackson.org/ERD/"

if ( Local_test ) {
  df_grp <- readRDS(paste0(Curated_path, "Block_group_data_AL.rds")) %>% 
    sf::st_as_sf()
  df_blk <- readRDS(paste0(Curated_path, "Block_data_AL.rds")) %>% 
    sf::st_as_sf()
} else {
  z <- url(paste0(DataLocation, "UCS_sfdata.rds"), method="libcurl")
  df_grp <- readRDS(z)
  close(z)
}

Depths <- as.character((0:20)/2)

States <- c("Alabama", "Florida", "Georgia", "Louisiana", "South Carolina",
            "Texas")
State_abbr <- c("AL", "FL", "GA", "LA", "SC", "TX")

State_df <- as_tibble(cbind(States, State_abbr))

#   Round up to the nearest 5
Max_poverty <- 5*round((max(df_grp$Pct_poverty)+2.5)/5)

# basemap <- leaflet() %>%
#   # add tiles
#   addProviderTiles("OpenStreetMap")
```

# {.sidebar}

```{r}

selectInput('State', 'Choose a state', States)

br()

sliderInput('Pct_poverty', '> % in Poverty', 
            min=5, max=Max_poverty,
            # min=5, max=100,
            value=10, 
            step=5, round=0)
br()

selectInput('Depth', 'Sea Level Rise (ft)', Depths)

br()

#   choose variable to color blocks with
radioButtons("Block_var", "Color Blocks by:",
             c("Population"="Pop_blk",
               "Avg Flood Depth"="Wmean",
               "Households"="Fam_house_blk",
               "% Block Flooded"="Flood_pct"))
# 
# checkboxInput('Show_blks', 'Show Blocks on mouse click', value=FALSE)

```

Data from NOAA Sea Level Rise datasets plus the 2020 ACS 
census data.

Note that Percent Poverty applies to those blocks at risk for a given value of
sea level rise - so for small rises in sea level and high poverty values, there
will be no blocks appearing.

Blocks are turned off by default. Use the layer tool in the upper left corner
to control which layer is visible.

Note that the visible data is also shown as a table in the Data tab. Rows
selected in the data table will be highlighted on the map.

[Documentation](https://docs.google.com/document/d/1ZYLhbcxoM2DShvo-eewVF4UHyNHtJVhjPB6KRr6JFI4/edit?usp=sharing)

# Plot

```{r}
# tmapOutput('map')
leafletOutput('map')
```

# Data

```{r}
# tableOutput('data')
DT::dataTableOutput('data')
```

```{r}
#| context: server

#####   Load in a new state

df_grp <- reactive({
  if ( Local_test ) {
    i <- which(State_df$States==input$State)
  readRDS(paste0(Curated_path,
                 paste0("Block_group_data_",State_df$State_abbr[i] ,".rds"))) %>%
    sf::st_as_sf()
  }
})

df_blk <- reactive({
  if ( Local_test ) {
    i <- which(State_df$States==input$State)
  readRDS(paste0(Curated_path,
                 paste0("Block_data_",State_df$State_abbr[i] ,".rds"))) %>%
    sf::st_as_sf()
  }
})

#####   Subset data

dataset_grp <- reactive({
  foo <- df_grp() %>% filter(Depth==input$Depth) %>%
      filter(Pct_poverty>=input$Pct_poverty) #%>%
})

dataset_blk <- reactive({
  df_blk() %>% filter(Depth==input$Depth) %>%
    filter(GEOID %in% unlist(dataset_grp()[["Blocks"]]))
})

#####     end subsetting

###################################################################
  #     Debugging tool
  # observeEvent(input$map_shape_click, {
  #     click <- input$map_shape_click
  #     print(paste("--1--", unlist(dataset_grp()[["Blocks"]]))) 
  #     # print(paste("--2--", unlist(dataset_grp()[["Blocks"]]))
  # }) #END OBSERVE EVENT
  observeEvent(input$data_rows_selected, {
      rows <- input$data_rows_selected
      print(paste("--1--", rows))
      # print(paste("--2--", unlist(dataset_grp()[["Blocks"]]))
  }) #END OBSERVE EVENT
###################################################################

# output$map <- renderTmap({
output$map <- leaflet::renderLeaflet({
  Grp_data <- dataset_grp() %>% sf::st_drop_geometry()
  Blk_data <- dataset_blk() %>% sf::st_drop_geometry()
  # map <- basemap
  
  if (FALSE) { # show water
  #   Add water grid
  map <- map %>%
    addRasterImage(Crop_tif, colors = pal, opacity = 0.5) %>%
    addLegend(pal = pal, values = terra::values(Crop_tif),
    title = "Water Depth") 
  }
  #   Block groups and blocks
  # map <- map %>%
  map <- leaflet() %>% 
    addTiles() %>% 
    addPolygons(  # block groups
    data = sf::st_transform(dataset_grp(), crs=googlecrs),
    group = "Block Grps",
    opacity = 0.2,
    weight = 1,
    fillOpacity = 0.2,
    popup = paste(
      "Pop at risk:", Grp_data$Pop_at_risk, "<br>",
      "Homes at risk:", Grp_data$Fam_hou_at_risk, "<br>",
      "Avg flood depth (ft):", Grp_data$Mean_depth, "<br>",
      "% in poverty:", Grp_data$Pct_poverty
      )
  ) %>% 
    addPolygons( #      Blocks
    data = dataset_blk(),
    group = "Blocks",
    color = "purple",
    # opacity = 1,
    weight = 1,
    fillOpacity = 0.2,
    popup = paste(
      "Block Name:", Blk_data$NAME, "<br>",
      "Population:", Blk_data$Pop_blk, "<br>",
      "Households:", Blk_data$Fam_house_blk, "<br>",
      "Average Water Depth:", Blk_data$Wmean, "<br>",
      "Pct of Block flooded:", Blk_data$Flood_pct
  ) %>%
    addLayersControl(overlayGroups = c("Block Grps", "Blocks"),
                   options = layersControlOptions(collapsed=F)) %>%
    hideGroup("Blocks")
  
  if (length(rv_table$rows)>0) {
    print(paste("--X--", dataset_grp()[rv_table$rows,]))
    map <- map %>% 
    addPolylines(data=dataset_grp()[rv_table$rows,], 
                 color="red", 
                 weight=3)
  }
  
  # tmap::tmap_options(basemaps="OpenStreetMap")
  # # tmap::tmap_options(basemap.server = c("OpenStreetMap")) # tmap 4.0
  # tmap_mode("view")
  # 
  # print("--- here I am")
  # tm <-
  # tm_shape(dataset_grp()) +
  #   tm_polygons("Pct_poverty", alpha=0.3,
  #               popup.vars=c("Blk_grp",
  #                            "Pop_at_risk",
  #                            "Fam_hou_at_risk",
  #                            "Mean_depth",
  #                            "Pct_pop_at_risk",
  #                            "Pct_poverty")) +
  #     tm_dots(size=0.02) +
  # tm_shape(dataset_blk()) +
  #   # tm_polygons("Pop_blk", alpha=0.3,
  #   tm_polygons(input$Block_var, alpha=0.3, border.col="purple",
  #               popup.vars=c("Block Name"="NAME",
  #                            "Population"="Pop_blk",
  #                            "Households"="Fam_house_blk",
  #                            "Average Water Depth"="Wmean",
  #                            # "Pct_Pop_at_risk",
  #                            "Pct of Block flooded"="Flood_pct"))  
  # 
  # if (length(rv_table$rows)>0) {
  #   print(paste("--X--", dataset_grp()[rv_table$rows,]))
  #   tm <- tm +
  #   tm_shape(dataset_grp()[rv_table$rows,]) +
  #     tm_borders(col="red", lwd=3) 
  # }
  # 
  #   tmap_leaflet(tm, in.shiny=TRUE) %>%
  #   leaflet::hideGroup("dataset_blk()")
  
})

  # reactive values for map
  # rv_map <-reactiveValues(Clicks=list())
  # 
  # observeEvent(input$map_shape_click, {
  #     click <- input$map_shape_click
  #     # print(paste("--1--",str(click)))
  #     # rv_map$Clicks<-c(rv_map$Clicks, click$id)
  #     # rv_map$Clicks <- stringr::str_extract(rv_map$Clicks, "\\d+$")
  #   if (!input$Show_blks){ # only if in group mode
  #       rv_map$Clicks <- stringr::str_extract(click$id, "\\d+$")
  #       print(paste("--2--", rv_map$Clicks))
  #   }
  # }) #END OBSERVE EVENT

# output$data <- renderTable({
output$data <- DT::renderDataTable({
  dataset_grp() %>% sf::st_drop_geometry()} %>% 
    select(Blk_grp, Pop_at_risk, Fam_hou_at_risk, Mean_depth, 
           Pct_pop_at_risk, Pct_poverty),
  colnames=c("Block Grp", 
             "Pop at Risk", 
             "Homes at Risk", 
             "Mean flood depth", 
             "% Pop at risk",
             "% Poverty")
)

rv_table <-reactiveValues(rows=list())

observeEvent(input$data_rows_selected, ignoreNULL = FALSE, {
    rv_table$rows <- input$data_rows_selected
}) #END OBSERVE EVENT

```
